# Reproducible Research: Peer Assessment 1
==========================================

## Loading and preprocessing the data
```{r echo=TRUE}
## Load and preprocess the data 
myData =read.csv("activity.csv",header=T, stringsAsFactors=F)
# convert the date
myData$date <- as.Date(myData$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
# Aggregate by date
dailyTotals <- aggregate(steps ~ date, myData, sum)
# Histogram 
hist(dailyTotals$steps, main="Total number of steps taken each day", 
     xlab="Number of steps")
# Mean & median 
meanSteps <- mean(dailyTotals$steps)
medianSteps <- median(dailyTotals$steps)
```
- Mean = `r meanSteps`
- Median = `r medianSteps`

## What is the average daily activity pattern?
```{r echo=TRUE}
# Aggregate by interval 
intervalMeans <- aggregate(steps ~ interval, myData, mean)
# Time series plot
plot(intervalMeans$interval, intervalMeans$steps, type="l",
     xlab="5-minute interval", ylab="Average number steps")
# Interval with max steps 
indx <- which.max(intervalMeans$steps)
maxInterval <- intervalMeans[indx,]$interval
maxSteps <- intervalMeans[indx,]$steps
```
- Maximum interval `r maxInterval` with `r maxSteps` steps 

## Imputing missing values
```{r echo=TRUE}
# Total missing values
totalMissing <- sum(is.na(myData$steps))
```
- Total missing values `r totalMissing`
```{r}
#  Fill in missing values
missing <- which(is.na(myData$steps))
means <- sapply(myData[missing,"interval"], function(x) {
        intervalMeans[ intervalMeans$interval==x, "steps"]
        })        
myData[missing, "steps"] <- means
# Histrogram
dailyTotals <- aggregate(steps ~ date, myData, sum)
hist(dailyTotals$steps, main="Total number of steps taken each day - Imputed NA's", 
     xlab="Number of steps")
# Mean & median 
naMeanSteps <- mean(dailyTotals$steps)
naMedianSteps <- median(dailyTotals$steps)
```
- Mean = `r naMeanSteps`
- Median = `r naMedianSteps`  
These values do not differ much from the dataset with missing values included 
so they have very little effect.


## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
# Weekday or weekend factor 
myData$week <- weekdays(myData$date)
myData$week <- sapply(myData$week, function(x) {
        if (x=="Sunday"|x=="Saturday"){
                "weekend"
        } else {
                "weekday"
        }
        })
myData$week <- factor(myData$week)
# Aggregate by interval for each: weekday & weekend
intervalMeans = aggregate(steps ~ interval + week, myData, mean)
# Time series plot
library(lattice)
xyplot(steps ~ interval | week, intervalMeans, type = "l", 
       layout = c(1, 2), 
       xlab = "5-minute Interval", 
       ylab = "Averag number of steps taken")

```

